---
title: "empirical"
author: "Lorraine Jiang"
date: "2/25/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Question 1a
```{r}
getwd()
library(KernSmooth)
library(tidyverse)

#1a
#read table
andro2 <- read.table("andro2.txt", header = TRUE)
signal <- andro2$Signal
t <- andro2$Time
# 35 points
points_35 <- subset(signal, between(t, 191.5, 208.5))
# mean sd variance of the 35 points
mean_35<- mean(points_35)
mean_35
sd_35 <- sd(points_35)
sd_35
var_35 <- sd_35^2
var_35
# value of the mean function
ssquared35 <- 1/34*sum((points_35-mean_35)^2)
ssquared35

# 105 points, we conduct the same procedure as the first did
points_105 <- subset(signal, between(t, 174, 226))
# mean sd variance of the 105 points
mean_105 <-mean(points_105)
sd_105<-sd(points_105)
var_105 <- sd_105^2
mean_105 
sd_105
var_105

ssquared105 <- 1/104*sum((points_105-mean_105)^2)
ssquared105

#95% CI for 35 points, we lose one degrees of freedom
t1<- qt(0.975,34) #t = 2.032245
mean_35 - (t1*sd_35/sqrt(35)) #lower
mean_35 + (t1*sd_35/sqrt(35)) #upper
#95% CI for 105 points
t2<- qt(0.975,104) #t= 1.983038
mean_105 - (t2*sd_105/sqrt(104)) #lower
mean_105 + (t2*sd_105/sqrt(104)) #upper
```
The value of the mean function at time 200 by first averaging the 35 points and 105 points is 0.01035496 and 0.00514793 respectively.The 95% interval for the 35 points and 105 points are [-0.01238412,0.05752698]&[-0.04499943, -0.01709581].

(b). The bias is not accounted for in my 95% confidence intervals from part(a) because the my confidence interval only explains random error, nothing to do with bias. Specifically, a larger variable will lead to a wider CI and larger margin of error; a smaller variable will cause a tighter CI and smaller margin of error.


(c)
```{r}
# we chose the bandwidth = 2
signal <- andro2$Signal
time <- andro2$Time
nwfit <- function(t, x, y, bw){
  GausK <- dnorm(x, mean = t, sd = bw)
  sum(y*GausK)/sum(GausK)
}
nw.hat <- sapply(time, nwfit, x = time, y = signal, bw = 2)
plot(time, signal, cex = 0.5)
lines(time, nw.hat, type = "l", col = "red", ylab = "NW")

```

(d)
```{r}
plot(time, signal, cex = 0.5)
loc.fit1 <- locpoly(x=time, y=signal, degree = 1, bandwidth = 2)
loc.fit2 <- locpoly(x=time, y=signal, degree = 1, bandwidth = 3)
loc.fit3 <- locpoly(x=time, y=signal, degree = 1, bandwidth = 5)
lines(loc.fit1$x, loc.fit1$y, type = "l", col = "red", ylab = "local regression")
lines(loc.fit2$x, loc.fit2$y, type = "l", col = "green", ylab = "local regression")
lines(loc.fit3$x, loc.fit3$y, type = "l", col = "blue", ylab = "local regression")
```

(e)
```{r}
library(IDPmisc)
peak1 <- peaks(loc.fit1$x, loc.fit1$y)
p1<- mean(diff(peak1$x[-1]))
peak2 <- peaks(loc.fit2$x, loc.fit2$y)
p2 <- mean(diff(peak2$x))
peak3 <- peaks(loc.fit3$x, loc.fit3$y)
p3 <- mean(diff(peak3$x[-1]))
p1
p2
p3
```
The average amount of time between the peaks is 69.43125

Question 2(a)
```{r}
#read table
GermanProduction <- read_table2("GermProd.txt")
#extract month and production separately
month <- GermanProduction$month
production <- GermanProduction$production
# fit a line by Nadaraya–Watson estimator
nw.hat <- sapply(month, nwfit, x = month, y = production, bw = 24)
ggplot(data = GermanProduction, aes(x=month, y=production)) + geom_point() +
  geom_line(aes(month, nw.hat, color = "NW")) ->p1
```

(b)\
NW estimator is given by
\

The red curve deviates to the data in the head and tails, where exist bias at the boundary. According to the formula, the NW estimator is inconsistent at the boundaries as observations are not corresponding to the estimation one by one.

(c)
```{r}
# Local Regression Fit
loc.fit <- locpoly(month, production, degree = 1, bandwidth = 24)
ggplot(GermanProduction, aes(x = month, y = production)) +
  geom_point() +
  geom_line(aes(month, nw.hat, color = "NW"), size = 1) +
  geom_line(data = as.data.frame(loc.fit), aes(x, y, color = "local-linear"))

loc.fit2 <- locpoly(month, production, degree = 2, bandwidth = 24)
p2 = p1 + geom_line(aes(month, loc.fit$y[1:400], color = "Linear")) +
  geom_line(aes(month, loc.fit2$y[1:400], color = "Poly"))
p2

```

(d)\
We have smaller bias at the boundary because of the adjustment. As we increase the order and degree, the curve better fitted the data，making the less bias. Thus, the local linear model minimizes the weighted least squares.

Question 3 (a)
```{r}
#read csv
WBushApproval <- read_csv("WBushApproval.csv")
WBushApproval$Date <- as.Date(WBushApproval$Start,format="%m/%d/%Y")



date <- WBushApproval$Date
approve <- WBushApproval$Approve
loc.fit <- locpoly(as.numeric(date), approve, bandwidth = 10, degree = 1)
ggplot(WBushApproval,aes(x = Date, y = Approve)) +
  geom_point() +
  geom_line(aes(as.Date(x, "1970-01-01"), y, colour = "local - linear"),
            data = as.data.frame(loc.fit), size = 1)


xnew <- as.Date("2005-06-01")
approx(loc.fit$x, loc.fit$y, xout = as.numeric(xnew))


nw.hat <- sapply(as.numeric(date), nwfit, x = as.numeric(date), y = approve, bw = 10)
xnew <- as.Date("2001-10-01")
# local linear estimator
approx(loc.fit$x, loc.fit$y, xout = as.numeric(xnew))
approx(loc.fit$x, loc.fit$y, xout = as.numeric(xnew))


ggplot(WBushApproval,aes(x = Date, y = Approve)) +
  geom_point() +
  geom_line(aes(date, nw.hat, color = "NW"), size = 1)

```
